### 2024-01-10 Manuscript_Nater_etal.qmd
Added "skeleton text" for second chapter of Discussion
Here, too, plenty to add

---
title: "Large-scale spatiotemporal variation in vital rates and population dynamics of an alpine bird"
format: html
author: Chloé R. Nater, James A. Martin, Erlend B. Nilsen
editor: visual
bibliography: rypeIDSM_MS2.bib
---

# Abstract

TBA

# Introduction

<!-- @JAMES & ERLEND: First draft is relatively "compact". Input on extending it and/or addition of relevant references appreciated -->

There is growing demand for biodiversity indicators from international unions, national government, local management bodies, and corporate and industry actors. Such indicators ideally represent a wide range of biodiversity's states and functions [e.g. Essential Biodiversity Variables, @pereira2013; @jetz2019], yet the development of suitable indicators for certain attributes, such as species abundance and demography, has been more difficult than for others [@schmeller2018; @waldock2022]. This is at least partially due to challenging requirements regarding spatial scales of useful biodiversity indicators. On one hand, indicators need to be representative at large geographic scales, for example, in the context of countries' reporting towards biodiversity targets [e.g. @feld2009]. On the other hand, indicators also ideally have good spatial resolution as the scales relevant for local-level management and planning are often much smaller [@stevenson2021]. This latter requirement is particularly crucial for (infrastructure) development strategies and for species management and conservation, both of which tend to require knowledge on species abundance and population dynamics (i.e. demographic rates) that is relevant for their county- or municipality-level decision making (+ refs?). Another reason why abundance and population indicators ideally come with good spatial resolution is that there can be substantial amounts of variation in population dynamics and life history of species across space [e.g. @robinson2014; @horswill2019]. This variation that needs to be accounted for to develop successful and sustainable strategies for area use, (harvest) management, and species and biodiversity conservation [@williams2002].

While large-scale, spatially-explicit indicators for species abundance and populations are clearly needed, development and practical implementation are greatly limited due to the reliance of such indicators on the availability of data from large-scale, long-term monitoring programmes [@proencca2017]. Consequently, many countries have been working on setting up, maintaining, and improving such monitoring programmes over the last decades. Many now well-established programmes focus on breeding birds and butterflies, and examples include the North American Breeding Bird Survey (<https://www.pwrc.usgs.gov/bbs/>), UK Butterfly Monitoring Scheme (<https://ukbms.org/>), the Game and Wildlife Conservation Trust Partridge Count Scheme [@aebischer2010], and the Swiss Biodiversity Monitoring (<https://www.biodiversitymonitoring.ch/>). <!-- is there any other one we want to name here specifically? -->

There is a natural trade-off between quality and quantity of data that can be collected in any monitoring programme: collecting high quality data in a structured manner is costly and requires trained specialists, thus limiting the amount of data that can be collected, while participatory monitoring, i.e. the collection of ecological data by members of the public [also called citizen or community science, @fraisl2022], allows to greatly reduce costs and extend spatial and taxonomic scales of monitoring at the expense of data quality and risk of bias [@johnston2023]. Consequently, many large-scale monitoring programmes are often limited to presence-absence or very simple count observations, making them suitable for the development of indicators of species distributions and perhaps population trends, but usually not of abundance, population dynamics, and demographic rates (+ example refs). The exception here are monitoring programmes that succeed in making use of a large number of citizen scientists that have been trained to collect data and record metadata in a structured manner and according to a carefully designed protocol. For example, in the United States hunters participate in the collection of bands and wings from harvested American Woodcock (*Scopolax minor)* to estimate survival and age ratios [@zimmerman2010]. <!-- can we think of 1-2 internationally well-known examples? Check out MAPS (Monitoring of Avian Productivity and Survivorship) and CES (Constant Effort Sites) --> In Norway there is a monitoring programme for terrestrial game bird species called "Hønsefuglportalen" (= "game bird portal", <https://honsefugl.nina.no/Innsyn/en>). It is a line transect survey programme carried out annually in \>120 localities across the country (\>2000 transects) by trained volunteers using pointing dogs. The programme has a well developed protocol for recording bird observations, auxiliary data, and relevant metadata (+ref to protocol?) and established routines for quality control and annual releases of publicly available data via GBIF. As such, it is particularly well suited to become part of a workflow for producing and updating abundance and population indicators on an annual basis.

The line transect data from "Hønsefuglportalen" has been used previously for estimating abundance trends of willow ptarmigan (*Lagopus lagopus*) across Norway [e.g. @bowler2020; @nilsen2020] but large-scale estimation of demographic rates underlying these trends has thus far remained as untapped potential of the dataset. @nilsen2023 recently developed a novel integrated distance sampling model (IDSM) which successfully uses the age of idividuals detected along line transects data coupled with radio-telemetry data to estimate not just abundance, but also survival and recruitment across years. In this study, we adapt the model of @nilsen2023 to run not just on a single site but on all areas with publicly available line transect data from "Hønsefuglportalen" simultaneously. Unlike several previous studies applying integrated models for population dynamics to multiple (sub-) populations separately and comparing results [e.g. @robinson2014; @nater2023], we opt for an approach explicitly integrating across space, thus allowing for sharing of information across locations and -- in effect -- space-for-time substitution [e.g. @horswill2019; @morrison2022]. We then apply the resulting multi-area IDSM to "Hønsefuglportalen" data on willow ptarmigan to estimate population size, age-structure, survival, recruitment, and impacts of small rodent abundance across 41 reporting districts and 15 years (2007-2021) for this popular small-game species. We further embed the modelling workflow in a reproducible, semi-automated pipeline that will greatly facilitate the repeated calculation of abundance and population indicators at different spatial scales as new data are added every year.

# Methods

## Study species

The willow ptarmigan is a tetraoid bird with a circumpolar distribution, mainly inhabiting sub-alpine and arctic ecosystems [see e.g. @fuglei2020]. While the species is currently listed as Least Concern (LC) both in the global and Norwegian Red List of Species, it has undergone rather dramatic declines in abundance in Norway since the turn of the 20th century [@hjeljord2022]. While the main reason for the long-term decline in abundance remain unresolved, willow ptarmigan are considered sentinel species that are sensitive to both climate change and land use changes [@Henden2017; @Storch2007]. Moreover, being one of only a handful of bird species that spend the winter in mountain ecosystems in Scandinavia, they are important components of the ecosystem as prey species for the gyrfalcon [@Alastair2020] <!-- do we have other "important" species to add here. Arctic fox? Some other endangered beasts perhaps? -->. The willow ptarmigan has a relatively fast pace of life [@sandercock2005; @Steen1996], and can display substantial spatio-temporal variation in demographic rates [@Bowler2020]. Their population dynamics are characterized by large inter-annual fluctuations in abundance [@hjeljord2022], and previous research has tied these fluctuations to rodent cycle through shared predators [@hagen1952; @Bowler2020]. This tight relationship between the breeding success of ground nesting birds and the rodent cycle is known as the Alternative Prey Hypothesis and has been a central part of Fennoscandian research on grouse population dynamics for many decades [@hagen1952; @Linden1988; @Steen1985]. In addition, spring weather conditions and phenology is known to have considerable effects on breeding success and recruitment rates [@Eriksen2023; @Steen1985]. Across their distributional range, they are iconic species with a high cultural value, partly linked to their popularity as game species. The latter means that information about spatio-temporal variation in demographic rates and population dynamics is particularly important in order to design sustainable harvest strategies. In addition, as a sentinel species the willow ptarmigan is well suited as an indicator species for ecosystem status; in Norway it is included in both the main national biodiversity [Nature Index for Norway, @jakobsson2020, <https://www.naturindeks.no/Indicators/lirype>] and ecosystem [Assessment of the Ecological Condition, @framstad2022] assessments.

## Data collection, management, and preparation

### Line transect sampling

The line transect survey data were collected through a structured participatory monitoring program called "Hønsefuglportalen" (<https://honsefugl.nina.no/Innsyn/en>). In August each year, trained volunteer fieldworkers collect observations of willow ptarmigan and other grouse species (rock ptarmigan *Lagopus muta*, black grouse *Lyrurus tetrix*, capercaillie *Tetrao urogallus*) along predefined line transects. To increase the detection probability, each team of fieldworkers use a pointing dog to locate the birds. Each survey team typically consist of at least two persons (one dog handler and one person responsible for following the transect line) and one dog. Often, more than one dog is used for a survey, but only one dog should be used at a time. The transect lines vary in length, but are typically between 1 and 8 km (range: 0.3-16.2 km, median: 3 km). When birds are observed, the exact location of observation is noted, along with its perpendicular distance from the transect line, as well as the age and sex of the birds. An observation typically includes 1 - 12 birds (mean = 5.6), with groups \> 1 typically representing one brood (female and or male with young-of-the-year chicks). When the surveys are conducted in August, the chicks of the year are able to fly but can be distinguished from older birds as they are still of smaller body size. Since 2019, most of the data has been collected using a mobile app tailored made for the monitoring program, where the field workers can register and get access to the transect lines allocated to them by the local organizers. Prior to 2019, data were collected on a dedicated fieldwork form, and entered manually in a web portal afterwards. After field data has been registered, it undergoes several steps of quality control carried out by local stakeholders and personnel from the Norwegian Institute for Nature Research (NINA). Surveys are carried out on both public and private land. After an initial embargo period, all data from public land are published and made freely available as a sampling-event data set on GBIF (refs). The published datasets contain both metadata about the transect surveys (survey date, line transect length and location, study area ID, etc.) as well as bird observation data (species, number of birds of different categories (adult males, adult females, juveniles, and birds of unknown category), perpendicular distance to transect line, exact location, and time of observation). Formally, the data from public land is published as three distinct data sets, one for each of the main public land administrators (Statskog, FeFo and Fjellstyrene, respectively). <!-- I changed "owners" to "administrators", just because to me "public" implies it is not owned by anyone. But if this is wrong, please change it back -->

Notably, the program is not designed as a centralized national monitoring programme, but rather a collection of local and regional survey programs. All involved survey areas use a common field protocol and data collection model. In addition, the local study designs are reviewed by staff at NINA, and common recommendations for study design are provided. However, because participation by stakeholders is voluntary, the spatial distribution of transects and sampling effort is not homogeneous. In general, the effort is higher in South-Eastern and Central Norway, medium in Northern Norway, and low in Western and Southern Norway.

In this study we used all publicly available data for the period 2007-2021, which included a total of 2225 transects in 41 different reporting districts spanning 9 counties and 50 municipalities. Transects on which no willow ptarmigan were ever observed (i.e. species absence likely due to low habitat suitability) were dropped, resulting in 2077 transects used in the analyses.

### Radio-telemetry study in Lierne

The model of @nilsen2023 integrated line transect data with radio-telemetry data from from an ongoing field study of marked willow ptarmigans in Lierne municipality in Central Norway. From 2015 to 2019, around 50 birds were captured in winter (late February or early March) each year and fitted with VHF collars. The marked birds were then monitored on a regular basis until they either i) died, ii) their transmitter's battery stopped working, or iii) we lost contact with the bird for other reasons. For most of the year, the birds were monitored <!-- how were they actually monitored? Can we have a tiny bit more detail here? --> at least once a month. During the breeding and chick-rearing season (May to July) birds were monitored more often, and during December and January we obtained few observations due to challenging field work conditions. A proportion of the birds were harvest annually in the regular recreational harvest, and birds that were harvested were reported as shot to the field personnel. In addition, as the collars had mortality switch, we were also able to locate and retrieve a high proportion of birds that died for natural causes, resulting in a known-fate mark-recapture dataset. The radio-telemetry study is described in detail in @Israelsen2020 and in @Arnekleiv2022.

In this study we used data from years 2015 - 2020, and the total sample size across these years was 139 birds for the Aug-Jan period and 258 birds for the Feb-Jul period.

## National-scale integrated model

### Integrated distance sampling model (IDSM) for willow ptarmigan

@nilsen2023 recently developed an integrated distance sampling model (IDSM) which jointly analyses line transect and radio-telemetry data and applied it to willow ptarmigan in the Western part of Lierne municipality in Norway. The model consists of a deterministic population model with two age classes (juveniles and adults) and four data likelihoods: 1) likelihood for observation distances from transect lines for estimating detection probability; 2) likelihood for age-specific counts on transect surveys for estimating numbers of juveniles and adults present; 3) likelihood for juvenile to adult ratios observed at the locality level to provide estimate recruitment rate (as juveniles/adult); and 4) likelihood for known-fate telemetry data to estimate seasonal and annual survival. Below, we describe our new extension of this model to include data from several as opposed to just one. For more detailed information on the single-site model, including tests of model performance, see @nilsen2023.

### Multi-area model extension

For applying the ptarmigan IDSM across all 41 reporting districts we included an area index in all model parameters (@fig-model) and enabled sharing of information among areas by explicitly modelling spatial variation alongside shared temporal and residual variation in vital rates and detection parameters.

The spatially-explicit formulation of the two age-class population model can be written as:

$$
\begin{aligned}
  D_{juv, x, j, t+1} & = D_{ad, x, j, t+1} * R_{x, t+1} \\
  D_{ad, x, j, t+1} & = S_{x, t} * (D_{juv, x, j, t} + D_{ad, x, j, t})  
\end{aligned}
$$

Here, $D_{juv, j, x, t}$ and $D_{ad, j, x, t}$ are the densities of juvenile and adult ptarmigan in survey site (= transect) $j$ of area $x$ in year $t$, respectively. Both juveniles and adult survive from year $t$ to $t+1$ with an area- ($x$) and year- ($t$) specific survival probability $S_{x, t}$, and survivors produce the next generation of juveniles according to an area- and year-specific recruitment rate ($R_{x,t}$).

![Simplified graphical representation of the ptarmigan life cycle with two age classes and the data sources included in the integrated distance sampling model. The pink "x"'s indicate the added dimension for area. Juv\[x,t\] = juveniles in year t. Ad\[x,t\] = adults in year t. R\[x,t\] = recruitment rate in year t. S\[x,t\] = survival probability from year t to t+1. Note that the additional site (=transect) dimension, "j", is omitted for the sake of illustration.](Figures/RypeIDSM_multiArea_structure.png){#fig-model}

The initial densities, $D_{juv, x, j, 1}$ and $D_{ad, x, j, 1}$ are modelled for each site (= transect) as random realizations of log normal distributions with area-specific log means ($\mu^{D1}_x$) and log standard deviations ($\sigma^{D1}_x$). Survival ($S_{x,t}$) and recruitment ($R_{x,t}$), on the other hand, are assumed to be the same for all sites $j$ within a given area $x$ and were modelled as:

$$
\begin{aligned}
  logit(S_{x, t}) & = logit(\mu^S) + \varepsilon^{X.S}_{x} + \varepsilon^{T.S}_{t} + \varepsilon^{R.S}_{x,t} \\
    log(R_{x, t}) & = log(\mu^R) + \beta_x*rodentOcc_{x,t} +  \varepsilon^{X.R}_{x} + \varepsilon^{T.R}_{t} + \varepsilon^{R.R}_{x,t}
\end{aligned}
$$ The global means ,$\mu$, and the normally distributed spatial random effects, $\varepsilon^X$, represent the equivalent of what is elsewhere referred to as "hyper-parameter distributions" for sharing information on demographic rates across areas [e.g. @horswill2019; @horswill2021]. We also used this same approach for defining the area-specific effects ($\beta_x$) of local yearly rodent occupancy ($rodentOcc_{x,t}$) on recruitment. In addition to spatial variation in survival and recruitment, we also included large-scale temporal variation through random year effects that were shared by across all areas ($\varepsilon^T_t$) and otherwise unaccounted for variation through year- and area-specific residual random effects ($\varepsilon^R_{x,t}$). Spatial, temporal, and residual random effects were modelled as normally distributed with globally defined (= shared) standard deviations.

The three likelihoods for data resulting from the line transect sampling (observation distances, age-specific counts, and juvenile to adult ratios, see above) were also formulated as spatially explicit, with year- and area-specific distance sampling detection parameters modelled in the same way as survival and recruitment (bar the effect of rodent occupancy, @fig-model). For the known-fate telemetry data (and the seasonal decomposition of survival estimated from it), on the other hand, we did not add an additional area dimension as this data was only available for the Lierne area.

### Model implementation

We implemented our multi-area IDSM in a Bayesian framework using NIMBLE version 1.0.1 [@valpine2017] in R version 4.3.1 [@R2023]. For the likelihood for line transect observation distances we used a custom half-normal distribution developed by Michael Scroggie in the "nimbleDistance" package (https://github.com/scrogster/nimbleDistance). We used non-informative uniform priors for all parameters and including biologically sensible boundaries where possible. We simulated complete sets of initial values for all model nodes prior to model running and using pre-defined seeds to ensure reproducibility. Using NIMBLE's standard samplers, we then ran 5 MCMC chains of 150k iterations each. We discarded the first 75k samples of each chain as burn-in, and thinned the remainder by a factor 25, resulting in a final joint posterior containing a total of 5 x 3k = 15k samples (note that high thinning rates were necessary to constrain memory load of the joint posterior, which included 314568 monitored parameters).

## Post-hoc variance decomposition

Following model fitting, we calculated posterior distributions for the proportions of variance in survival probabilities, recruitment rates, and detection decay explained by 1) spatial variation ($var_{area}$), 2) temporal variation ($var_{year}$), 3) residual variation ($var_{residual}$), and 4) variation in rodent occupancy ($var_{rodent}$). To obtain the proportion variance explained by each of the component, we divided it by the sum of all other components ($var_{area} + var_{year} + var_{residual} + var_{rodent}$). The spatial, temporal, and residual variance components were defined as the square of the estimated corresponding random effects standard deviation from the model while $var_{rodent}$ was calculated as the variance of all area- and year-specific $\beta_x*rodentOcc_{x,t}$ products.

## Reproducible workflow with "targets"

Reproducibility and ease of repeating analyses was a key focus when developing the multi-area IDSM.To that end, we set up the workflow as a "targets pipeline", implemented through the R package "targets" [@landau2021]. The pipeline contains a variety of options for controlling modelling decisions in the workflow such as the year range of data to consider, the level of spatial aggregation (i.e. reporting district vs. survey locality), whether to model time variation in survival and/or effects of rodent abundance, whether to run MCMC chains sequentially or in parallel, etc. A visual representation of the pipeline is also displayed in @fig-workflow and for more details on pipeline implementation and options, we refer the reader to the GitHub repository: <https://github.com/ErlendNilsen/OpenPop_Integrated_DistSamp>.

![Graphical representation of the "targets pipeline" for the multi-area modelling setup. Upward facing triangles are functions, downward facing triangles are general options/arguments, circles are objects and outputs (="targets") created as part of the workflow.](Figures/Workflow_layout.png){#fig-workflow width="424"}

# Results

```{r, echo = FALSE}
# Posterior summaries
post.sum <- readRDS("Results/PosteriorSummaries_byArea.rds")

# Printing function
printPostSum <- function(table, row = NULL, area = NULL, exp = FALSE, pCalc = FALSE){

  if(!missing(area)){
    row <- which(table$Area == area)
  }

  if(!("Area" %in% colnames(table))){
    table$Area <- "None"
  }

  if(exp){
    table$Median <- exp(table$Median)
    table$lCI <- exp(table$lCI)
    table$uCI <- exp(table$uCI)
  }

  if(pCalc){
     pi <- 3.141593
     W <- 200
     table$Median <- sqrt((pi*table$Median^2)/2)/W
     table$lCI <- sqrt((pi*table$lCI^2)/2)/W
     table$uCI <- sqrt((pi*table$uCI^2)/2)/W
  }

  print.out <- list(
    area.name = table$Area[row],
    post.sum = paste0(
      round(table$Median[row], digits = 3), 
      " [",
      round(table$lCI[row], digits = 3),
      ", ",
      round(table$uCI[row], digits = 3),
      "]"
    )
  )
  return(print.out)
}

# Detection stats
Det.hyper.mean <- printPostSum(table = post.sum$hParams.sum, 
                              row = which(post.sum$hParams.sum$Parameter == "h.mu.dd"),
                              exp = TRUE)
pDet.hyper.mean <- printPostSum(table = post.sum$hParams.sum, 
                              row = which(post.sum$hParams.sum$Parameter == "h.mu.dd"),
                              exp = TRUE, pCalc = TRUE)

Det.sd.spatial <- printPostSum(table = post.sum$hParams.sum,
                             row = which(post.sum$hParams.sum$Parameter == "h.sigma.dd"))
Det.sd.temporal <- printPostSum(table = post.sum$hParams.sum,
                             row = which(post.sum$hParams.sum$Parameter == "sigmaT.dd"))
Det.sd.residual <- printPostSum(table = post.sum$hParams.sum,
                             row = which(post.sum$hParams.sum$Parameter == "sigmaR.dd"))
Det.min <- printPostSum(table = post.sum$detect.sum, 
                      row = which(post.sum$detect.sum$Median == min(post.sum$detect.sum$Median)))

pDet.min <- printPostSum(table = post.sum$detect.sum, 
                      row = which(post.sum$detect.sum$Median == min(post.sum$detect.sum$Median)),
                      pCalc = TRUE)

Det.max <- printPostSum(table = post.sum$detect.sum,
                      row = which(post.sum$detect.sum$Median == max(post.sum$detect.sum$Median)))

pDet.max <- printPostSum(table = post.sum$detect.sum,
                      row = which(post.sum$detect.sum$Median == max(post.sum$detect.sum$Median)),
                      pCalc = TRUE)

# Survival stats
S.hyper.mean <- printPostSum(table = post.sum$hParams.sum, 
                              row = which(post.sum$hParams.sum$Parameter == "h.Mu.S"))
S.sd.spatial <- printPostSum(table = post.sum$hParams.sum,
                             row = which(post.sum$hParams.sum$Parameter == "h.sigma.S"))
S.sd.temporal <- printPostSum(table = post.sum$hParams.sum,
                             row = which(post.sum$hParams.sum$Parameter == "sigmaT.S"))
S.sd.residual <- printPostSum(table = post.sum$hParams.sum,
                             row = which(post.sum$hParams.sum$Parameter == "sigmaR.S"))
S.min <- printPostSum(table = post.sum$pSurv.sum, 
                      row = which(post.sum$pSurv.sum$Median == min(post.sum$pSurv.sum$Median)))
S.max <- printPostSum(table = post.sum$pSurv.sum,
                      row = which(post.sum$pSurv.sum$Median == max(post.sum$pSurv.sum$Median)))

# Recruitment stats
R.hyper.mean <- printPostSum(table = post.sum$hParams.sum,
                             row = which(post.sum$hParams.sum$Parameter == "h.Mu.R"))
R.sd.spatial <- printPostSum(table = post.sum$hParams.sum,
                             row = which(post.sum$hParams.sum$Parameter == "h.sigma.R"))
R.sd.temporal <- printPostSum(table = post.sum$hParams.sum,
                             row = which(post.sum$hParams.sum$Parameter == "sigmaT.R"))
R.sd.residual <- printPostSum(table = post.sum$hParams.sum,
                             row = which(post.sum$hParams.sum$Parameter == "sigmaR.R"))
R.min <- printPostSum(table = post.sum$rRep.sum, 
                      row = which(post.sum$rRep.sum$Median == min(post.sum$rRep.sum$Median)))
R.max <- printPostSum(table = post.sum$rRep.sum,
                      row = which(post.sum$rRep.sum$Median == max(post.sum$rRep.sum$Median)))

# Rodent effects stats
betaR.hyper.mean <- printPostSum(table = post.sum$hParams.sum,
                             row = which(post.sum$hParams.sum$Parameter == "h.Mu.betaR.R"))
betaR.sd.spatial <- printPostSum(table = post.sum$hParams.sum,
                             row = which(post.sum$hParams.sum$Parameter == "h.sigma.betaR.R"))
betaR.min <- printPostSum(table = post.sum$betaR.sum, 
                       row = which(post.sum$betaR.sum$Median == min(post.sum$betaR.sum$Median)))
betaR.max <- printPostSum(table = post.sum$betaR.sum,
                      row = which(post.sum$betaR.sum$Median == max(post.sum$betaR.sum$Median)))
n.pos.betaR <- length(which(post.sum$betaR.sum$Median > 0))
n.neg.betaR <- length(which(post.sum$betaR.sum$Median < 0))

# Population density stats
popDens.sum <- subset(post.sum$popDens.sum, SummaryPeriod == "2018-2021")
dens.min <- printPostSum(table = popDens.sum, 
                      row = which(popDens.sum$Median == min(popDens.sum$Median)))
dens.max <- printPostSum(table = popDens.sum,
                      row = which(popDens.sum$Median == max(popDens.sum$Median)))

# Population growth rate stats
lambda.sum <- subset(post.sum$lambda.sum, SummaryPeriod == "2018-2021")
lambda.min <- printPostSum(table = lambda.sum, 
                      row = which(lambda.sum$Median == min(lambda.sum$Median)))
lambda.max <- printPostSum(table = lambda.sum,
                      row = which(lambda.sum$Median == max(lambda.sum$Median)))
n.pos.lambda <- length(which(lambda.sum$Median > 1))
n.neg.lambda <- length(which(lambda.sum$Median < 1))

```

All numerical results in the following are presented as median \[95% credible interval\] unless otherwise indicated. Posterior summaries (median, 95% credible interval, mean, standard deviation, coefficient of variation) for all main parameters are also provided in the supplementary file "PosteriorSummaries_byAreas.csv". Supplementary figures (SF) are provided as .pdf files with captions in "SuppFigures_Captions.txt".

## Population density

Only during the most recent four years (2018-2021) has data been collected regularly for all reporting areas included in the analyses. During this period, estimated population densities varied between `r dens.min$post.sum` birds/km^2^ in the area "`r dens.min$area.name`" close to the coast in central Norway to `r dens.max$post.sum` birds/km^2^ in "`r dens.max$area.name`" further south near the Swedish border. In general, population density appeared to be lowest in the Northern parts of Norway and highest in the eastern part of central Norway (@fig-densA). Uncertainty in density estimates was relatively consistent, with a few areas (inlcuding the one with the lowest estimated density, "`r dens.min$area.name`") sticking out by having substantially less precise estimates (@fig-densB). Populations fluctuated substantially over time in any given area (SF "TimeSeries_popDens1.pdf") and some years seemed to be indicative of relatively high (e.g. 2011, 2014, 2018) or low (e.g. 2012, 2015) densities across a large number of areas.

::: {#fig-dens layout-ncol="2"}
![Median density estimates](Figures/Avg_popDens_Map-4.png){#fig-densA}

![Uncertainty in density estimates](Figures/Avg_popDens_Map-6.png){#fig-densB}

Median (a) and coefficient of variation (standard deviation / mean) (b) of posterior estimates of average ptarmigan density in the four most recent years (2018-2021) across 41 reporting areas (summarised at the municipality level) in Norway. Darker colors indicate higher median values and higher uncertainty.
:::

## Population growth rate

Average population growth rates over the last four years (2018-2021) ranged from moderate declines (`r lambda.min$post.sum` in the "`r lambda.min$area.name`" area) to \> 50% increase (`r lambda.max$post.sum` in the "`r lambda.max$area.name`" area). In the majority of reporting areas (`r n.pos.lambda` out of 41), populations of willow ptarmigan have been increasing over the period 2018-2021 (@fig-lam). Some areas -- predominantly in central Norway -- also had declining populations, but many of those declines followed upon periods of increase between the start of data collection in 2007 and sometime between 2016 and 2018 (SF "TimeSeries_popDens1.pdf").

::: {#fig-lam layout-ncol="2"}
![Median population growth rate estimates](Figures/Avg_lambda_Map-4.png){#fig-lamA}

![Uncertainty in population growth rate estimates](Figures/Avg_lambda_Map-5.png){#fig-lamB}

Median (a) and standard deviation (b) of posterior estimates of average annual population growth rate over the four most recent years (2018-2021) across 41 reporting areas (summarised at the municipality level) in Norway. In a), pinkish colors indicate declining populations while greenish colors indicate growing populations (white = stable populations). In b) darker colors indicate higher uncertainty.
:::

The highest population growth rates were estimated for areas with relatively low population densities across latitudes but we did not find evidence for a strong association between population growth rates and population densities overall ([@fig-paramCorrs] A).

![Posterior summaries (points = posterior medians, lines = 95% credible intervals) of area-specific population growth rate vs. population density (A), recruitment rate vs. survival probability (B) and rodent effect on recruitment along a longitudinal gradient (C). Color indicates latitude of the midpoint of each area.](Figures/AllParams_Latitude.png){#fig-paramCorrs}

## Survival probabilities and recruitment rates

Annual survival probabilities ranged from `r S.min$post.sum` (area "`r S.min$area.name`") to `r S.max$post.sum` (area "`r S.max$area.name`") across reporting areas in Norway, with the highest values occurring in the far north and in the mountains in the south (@fig-pVRA). The global average survival probability across all areas and years ($\mu^S$) was estimated at `r S.hyper.mean$post.sum`. Spatial variation in survival (random effect SD = `r S.sd.spatial$post.sum`) was relatively low compared to temporal (`r S.sd.temporal$post.sum`) and residual (`r S.sd.residual$post.sum`) variation.

Recruitment rates varied between `r R.min$post.sum` (area "`r R.min$area.name`") and `r R.max$post.sum` (area "`r R.max$area.name`") and displayed a spatial pattern opposite to that of annual survival, i.e., lower recruitment rates co-occurring with higher survival rates and vice-versa (@fig-pVRB; @fig-paramCorrs B). Across all areas and years, average recruitment rate was `r R.hyper.mean$post.sum`. Unlike for survival, the model predicted similar magnitudes of spatial and temporal variation (random effect SDs of `r R.sd.spatial$post.sum` and `r R.sd.temporal$post.sum`, respectively), and about twice as much residual variation (`r R.sd.residual$post.sum`).

::: {#fig-pVR layout-ncol="2"}
![Median survival probabilities](Figures/Avg_pSurv_Map-1.png){#fig-pVRA}

![Median recruitment rates](Figures/Avg_rRep_Map-1.png){#fig-pVRB}

Posterior medians of average annual survival probabilities (a) and recruitment rates (b) across 41 reporting areas (summarised at the municipality level) in Norway. Darker colors indicate higher median values. Measures for corresponding uncertainty in estimates are visualized in SFs "Avg_pSurv_Map.pdf" and "Avg_rRep_Map.pdf" for survival and recruitment, respectively.
:::

The MCMC chains for many of the area-specific average survival probabilities and recruitment rates, as well as for the global averages (intercepts) for both vital rates, were mixing rather poorly. Despite that, mixing was good and resulting posteriors well defined for the area- and year-specific estimates of survival and recruitment (SF "PostDens_tS_tR.pdf"). There was substantial variation in both vital rates across time (SFs "TimeSeries_pSurv.pdf" and "TimeSeries_rRep.pdf"). In many areas, the years 2011, 2014, and 2018 not only sported high population densities (see above) but were also characterized by both high recruitment and low subsequent survival. The overall low density years 2012 and 2015, conversely, often featured lower recruitment and, in some cases, higher survival. Notably, there were also years with very little spatial synchrony, i.e. very different relative yearly survival probabilities and recruitment rates (e.g. 2010 and 2020 for survival and 2013, 2016, and 2017 for recruitment).

## Effects of rodent occupancy

The model predicted a positive global effect of rodent occupancy on recruitment rate (average slope on the log scale = `r betaR.hyper.mean$post.sum`). Nonetheless, spatial variation in the rodent effect was substantial (random effect SD = `r betaR.sd.spatial$post.sum`). This resulted in negative median effects in `r n.neg.betaR` areas, positive median effects in `r n.pos.betaR` areas, and a range of effect sizes from `r betaR.min$post.sum` (area "`r betaR.min$area.name`") to `r betaR.max$post.sum` (area "`r betaR.max$area.name`", @fig-paramCorrs C, SF "Rep_betaR.R.pdf"). The largest positive rodent effects were estimated for areas in the very North of Norway, as well as in the mountaineous regions in the central and southwestern parts of the country (SF "betaR_Map.pdf"). Effects with negative posterior medians were located mostly at intermediate latitudes, but we note that all of these had posterior distributions featuring substantial overlap with 0 (@fig-paramCorrs C).

## Detection parameters

Detection decay parameters, which determine detection probability in distance sampling surveys, varied across areas from between `r Det.min$post.sum` in "`r Det.min$area.name`" to `r Det.max$post.sum` in "`r Det.max$area.name`, resulting in detection probabilities over the transect sites (truncation distance = 200 m) from `r pDet.min$post.sum` to `r pDet.max$post.sum`, respectively. The global average detection decay was `r Det.hyper.mean$post.sum` (detection probability = `r pDet.hyper.mean$post.sum`), and in general, higher values were more common in the Southern half of the country than the Northern half (SF "Avg_detect_Map.pdf"). Variation in detection over time was modest (SF "TimeSeries_pDetect.pdf"), and among-year variation in detection decay (random effect log SD = `r Det.sd.temporal$post.sum`) was almost identical in magnitude to spatial (`r Det.sd.spatial$post.sum`) and residual (`r Det.sd.residual$post.sum`) variation.

## Variance decomposition

```{r, echo = FALSE}
# Posterior summaries
varDec.sum <- readRDS("Results/PosteriorSummaries_VarDecomp.rds")

# Printing function
printPostSum2 <- function(table, Parameter, VarComponent){

  idx <- which(table$Parameter == Parameter & table$VarComponent == VarComponent)

  print.out <- paste0(
      round(table$median[idx]*100, digits = 1), 
      " [",
      round(table$lCI[idx]*100, digits = 1),
      ", ",
      round(table$uCI[idx]*100, digits = 1),
      "]"
  )
  return(print.out)
}

# All parameter summaries
Rec_A <- printPostSum2(varDec.sum, "Recruitment", "Area")
Rec_Y <- printPostSum2(varDec.sum, "Recruitment", "Year")
Rec_R <- printPostSum2(varDec.sum, "Recruitment", "Residual")
Rec_Cov <- printPostSum2(varDec.sum, "Recruitment", "RodentOcc")

Surv_A <- printPostSum2(varDec.sum, "Survival", "Area")
Surv_Y <- printPostSum2(varDec.sum, "Survival", "Year")
Surv_R <- printPostSum2(varDec.sum, "Survival", "Residual")

Det_A <- printPostSum2(varDec.sum, "Detection", "Area")
Det_Y <- printPostSum2(varDec.sum, "Detection", "Year")
Det_R <- printPostSum2(varDec.sum, "Detection", "Residual")
```

The relative importance of different components for explaining parameter variation differed among recruitment rate, survival probability, and detection decay (@fig-varDecomp). The largest portion of variation in recruitment was attributed to residual variation (`r Rec_R` %), followed by spatial (`r Rec_A` %) and temporal (`r Rec_Y` %) variation. Rodent occupancy, which contains both a spatial and a temporal dimension, explained `r Rec_Cov` % of the total variation. For survival, there was large uncertainty in the estimated proportions of variance explained by different components. The model predicted similar potential contributions from temporal (`r Surv_Y` %) and residual (`r Surv_R` %) variation and suggested that spatial variation was only responsible for `r Surv_A` % of the total variance. Total variance in detection decay was attributed evenly to spatial, temporal, and residual variation at `r Det_A` %, `r Det_Y` %, and `r Det_R` %, respectively.

![Posterior distributions for the proportions of parameter variance explained by spatial (blue), temporal (orange), and residual (green) variation, as well as by effects of rodent occupancy (yellow).](Figures/VarDecomposition_overall.png){#fig-varDecomp}

# Discussion

Building on the work of @nilsen2023, we applied a novel integrated population model to data collected through a national-scale participatory monitoring programme to estimate spatial and temporal variation in demography of a culturally important game bird species, the willow ptarmigan. While our study was exploratory in nature, it recovered patterns consistent with ecological and life-history theory including trade-offs between survival and recruitment, and a tendency towards slower life histories at higher latitudes and altitudes. Space-for-time substitution also provided the statistical power necessary for the analysis to provide evidence for the alternative prey hypothesis, i.e. ptarmigans benefiting from high abundance of alternative prey for their predators. <!-- To me, this little paragraph has "abstract vibes", or maybe "conclusion vibes". I'll leave it here for now but note that it may get moved eventually.

## Abundance and vital rates across space and time

<!-- @ERLEND & @JAMES thoughts on what we see that is interesting, cool, expected, unexpected etc. Can we compare to other large scale studies, either on other grouse-like creatures or harvested species more generally? -->

<!-- Spatial perspective -->

The wide spatial distribution of the line transect monitoring afforded us the opportunity to explore variation in population density and vital rates across a relatively large spatial extent. Ptarmigan densities across the 41 reporting districts included in our analyses varied from around 2 birds/km^2^ to 55 birds/km^2^, with the lowest densities occurring far north in the country, as well as on the west coast and in the mountains in central Norway [@fig-densA]. These areas are characterized by more extreme climatic conditions, boasting cold temperatures and – in the North – a short growing season. Resulting reduced primary production limits food availability and as ptarmigan are income breeders using food resources acquired from nesting areas to supply energy and nutrients for egg production and incubation [@sandercock2005], lower carrying capacity in such areas is to be expected. <!-- @ERLEND do we have some more thoughts to add here? -->

We found increasing population trends over recent years in over half of the reporting districts [@fig-lamA]. <!-- relate this to changes in hunting management in recent years? Any other factors we think could have contributed? --> Nonetheless, recent population declines were also evident in some areas, and particularly in the mountains in central Norway [@fig-lamA]. <!-- What could be going on here? It's also interesting that while some of these areas have declines over the most recent years, that usually follows on a period of increase, or at least stability, up to around 2016 -->

While we found the highest population growth rates in areas with relatively low density, and relatively low growth rates in high-density areas [@fig-paramCorrs A], our analysis did not provide evidence for a strong relationship between density and growth rates across space. However, this is unsurprising as we did find substantial spatial differences in demography, which suggests that carrying capacity too varies over space. Overall, annual recruitment rates showed a spatial pattern that was inversely related to that of annual survival, with lower recruitment rates co-occurring with higher survival rates (and lower densities) and vice-versa [@fig-paramCorrs B], as we would expecte based on life history theory [@stearns1992]. Furthermore, the areas with relatively lower recruitment and higher survival were primarily found in the northern parts of Norway and in the mountainous region in the central part of the country, which aligns with previous studies finding relatively slower bird life histories in alpine / high altitude areas [e.g. @sandercock2005; @bears2009; @Wilson_Martin_2011; @alice2016]. <!-- What about the latitude? Do we know of any studies looking at latitudinal variation in life histories? When I last searched, I only found studies focusing on altitude -->

<!-- Temporal perspective -->
Population densities in all areas were subject to substantial variation across years, featuring periods of stability, increase, and decrease (SF "TimeSeries_PopDens1.pdf"). In most areas, there were also strong year-by-year fluctuations in population density on top of longer-term trends. Some of the resulting "high density years" were highly synchronized across large spatial scales, such as the years 2011, 2013, and 2018. Taking a closer look, we find that these are years that are characterized by high recruitment (SF "TimeSeries_rRep.pdf") and low subsequent survival (SF "TimeSeries_pSurv.pdf"), which often resulted in steep population declines towards the following year. The fact that these same years also match up with observed peaks in rodent abundance in many areas provides evidence for the Alternative Prey Hypothesis (APH). The AHP stipulates that high abundance of alternative prey (rodents, in this case) for common predators leads to population growth, and is well-supported throughout the literature for a range of taxa [e.g., @hagen1952; @Kjellander_Nordström_2003; @Reif_Tornberg_Jungell_Korpimäki_2001], including willow ptarmigan [@bowler2020]. <!-- Maybe another sentence or two about the low survival following the peak, which could be either due to negative density-dependence leading to resource competition among ptarmigan, or due to the rodent peak allowing high recruitment also for predators, which then in turn will have increased impacts on all prey populations in the subsequent year? -->

Taken together, this alludes to the large role small rodents play in shaping ptarmigan demography and population dynamics and is further supported by the largely positive effects of rodent abundance on recruitment estimated by our model (@fig-paramCorrs C). Once more, the highest latitude and highest altitude areas stood out, sporting the strongest rodent effects (SF "betaR_Map.pdf"). This may be linked to the overall low ptarmigan densities in these areas, as e.g. @Nielsen_1999 found evidence that gyrfalcon predation on ptarmigan was most impactful when ptarmigan densities were low. 
<!-- This previous sentence is adapted from a suggestion by James, but I feel like we need some more explanation/arguments here to clarify this. Perhaps also some other and/or more specific hypotheses= -->
Notably, the conclusive estimation of overall positive effects of rodent abundance on recruitment in our model was only possible thanks to the integration and sharing of data across multiple areas. When @nilsen2023 fit the IDSM to data from only a single area, they were unable to obtain a reliable estimate for the rodent effect due to limited statistical power. Consequently, the space-for-time substitution that comes with extending the model across mulitple area allows estimation of covariate effects that otherwise cannot be estimated, and opens up for future possibilities for studying effects of not just rodents, but also other environmental drivers on ptarmigan population dynamics. 
For example, [...] <!-- @ERLEND here we could add some thoughts on other covariates that may matter and are realistic/relevant/realistic to include in future studies. Best if it kind of naturally leads over to (and does not overlap too much with) the next paragraph  -->

<!-- Variation beyond rodents -->
Extending the present analysis by including additional covariates may also help with better understanding the mechanisms underlying the large portion of demographic rate variation that could only be attributed to random variation so far. This is the case especially for the relatively large residual variation [@fig-varDecomp] but also relevant for constant spatial and shared temporal variation. Interestingly, we found that a larger portion of variation in recruitment was allocated to spatial than to temporal factors, while the opposite was the case for survival. [...]
<!-- Any hypotheses/thoughts related to large- vs. local-scale drivers to explain this? -->

## Implications for management

Management decisions made at the resolution of large geo-political boundaries (e.g., Norway) run a high risk of being inadequate when there is substantial spatial variation in demographic processes and population dynamics, as is the case for willow ptarmigan. In Norway, willow ptarmigan -- and small game in general -- is managed [...] <!-- Brief summary of how and at what spatial scales patrmigan/small game management works in Norway -->
The results from ur study suggests that management decisions should be spatially refined and ideally informed by knowledge about the "local" population processes. [...] <!-- Some specific implications of the results for management of hunting and of habitat (e.g. better managed hunting is good because we do see increases in majority of local populations, but the mountain areas have had recent declines that may call for additional local measures, these measures may need to be adjusted since the mountain populations appear to have slower life histories and likely different sensitivities, etc.) -->

Besides pointers on how to adjust management of ptarmigan and their habitat, our results also provided some insights into the value, and possibly opportunities for improving the monitoring programme. First and foremost, our study demonstrates the tremendous potential lying in coordinating structured monitoring using a common protocol and training programme, as well as data processing pipeline. These were indeed the prerequisites that allowed us to easily and efficiently integrate data collected across the entire country in a joint analysis, and draw inference on spatio-temporal variation in demography and population dynamics at a large scale but also with good spatial resolution. 
While overall less variable across space and time than vital rates, differences in detection probabilities were nonetheless evident (SFs "Avg_detect_Map.pdf" and "TimeSeries_pDetect.pdf") and may help with mapping out potential for improvement in the monitoring programme. Particularly, we found generally lower detection probabilities in the northern half of Norway. [...] <!-- Some thoughts on why that might be. Could be related to landscape features and/or available observer pool. Any suggestions on what could be done to improve detection in the North, or at least "make up" for it? -->
Ultimately, increased detection probability would contribute to obtaining more precise estimates of both population density and demographic rates, which -- in turn -- would be of great value in particular in areas with relatively low population densities, low number of transects, and less years of data.

## Model limitations and outlook

-   Some limitations of "base model" that need mentioning (e.g. age assignment, ignoring potential effects of flock size on detection)

-   Outlook: improved modelling of spatio-temporal patterns through correlated REs

Another approach to maximising information gain when jointly analysing data from different areas is explicit modelling of spatial and spatio-temporal variation. While this is commonly done for modern species distribution models [e.g. @pacifici2017; @guelat2018], it has rarely been implemented in demographic models. This is not least due to the fact that few demographic models span large spatial extents. Since our IDSM and the willow ptarmigan data do actually do that though, extending the framework to account for spatial and spatiotemporal correlations may be feasible. Per now, we have not yet attempted this, but are planning to do so in the near future. (BUT: we do already have some computational limitations now, so likely to need development of more efficient sampling strategies. Which is fortunately possible with NIMBLE but there is a chance that another platform, e.g. INLA, may be required) <!-- @ERLEND: related to this, I want to discuss the potential of including the not publicly available data. Is there an overview somewhere over what the coverage (in terms of municipalities) would be if we had access to all the Hønsefuglportalen data? This might matter not just for increasing sample size, but perhaps also making CAR modelling (spatial structure through neighborhood relations) possible -->

<!-- @MYSELF: Have a closer look at chapter 19 in IPM book to see what they did there and also what other work they reference -->

-   Limited usefulness of auxiliary radio-telemetry data as only from one area (big potential for including additional datasets from other small-scale studies, telemetry but also other, e.g. nest monitoring?) <!-- @ERLEND: Any thoughts on what the status here is? Might there be other data available from beyond Lierne? Are there exciting future prospects e.g. linked to automated monitoring with cameras/sound boxes at a larger scale? -->

-   Outlook: integration with / coupling to models for harvest to develop tools for (automatic) calculation of e.g. hunting quotas <!-- @JAMES: This one is for you. Some ruminations on how and why it would be great to formally model harvest, integration with hunting data, maybe eventually develop a tool to calculate quotas (I know this has been developed for other game species), etc. -->

## Reproducible workflows for a sustainable future

-   Structure / setup of this workflow (from data collection to analytical results)

-   Why this matters: transparency and credibility, feedback to volunteer data collectors, cost-effective to re-do analyses for annual/regular reporting, easy to re-use and build upon, publicly accessible and well documented to allow more effective resource use in follow-up projects (sustainable use of funding), etc.

-   Also: useful to link up to different reporting routes and/or more general (multi-indicator) projects <!-- @EVERYONE: Do we know of some good references for this? Prseumably we cannot cite our chapter in the Technical Manual yet :-P -->

## Conclusion

-   Summarise neatly
-   Things worth highlighting: inference across large spatial scale but with "useful" resolution for management, advantages of integration across space (we can estimate better and more), integration of small locality-specific datasets with large national-scale monitoring, importance of reproducible workflows

# Acknowledgements

TBA ... Additionally, we would like to thank Bernardo Brandão Niebuhr dos Santos for his help with spatial data visualization. <!-- @ERLEND: you maybe know best all the people/institutions we need to thank here --> <!-- @JAMES: should we be acknowledging your fullbright scholarship which made the collab possible? If so, please add details. -->

# Data and code availability

The raw data from the line transect surveys is deposited on GBIF and can be accessed freely via the Living Norway Data Portal (https://data.livingnorway.no/). The work presented above is based on versions 1.7, 1.8, and 1.12 for the datasets from Fjellstyrene (key b49a2978-0e30-4748-a99f-9301d17ae119), Statskog (key 6a948a1c-7e23-4d99-b1c1-ec578d0d3159), and FeFo (key c47f13c1-7427-45a0-9f12-237aad351040), respectively.

The auxiliary radio-telemetry data, rodent occupancy data, and all code including the "targets pipeline" can be found in the project's repository on GitHub: <https://github.com/ErlendNilsen/OpenPop_Integrated_DistSamp>. The results presented in this paper were created using version 2.0 of the code (Zenodo citation TBA).

# References